---
title: "Subscription Preferences"
author: "Huizhong Peng"
date: "2024-11-07"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(dplyr)
library(ggplot2)
library(plotly)

options(scipen = 999)

knitr::opts_chunk$set(
        collapse = TRUE, 
        message = FALSE, 
        warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
  )

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
final = readRDS("final.rds")
users_data = read_csv("datasets/users_data.csv")
gyms_data = read_csv("datasets/gym_locations_data.csv")
```

## Brief intro

Among all the subscription plans, the monthly fee for basic plan is \$19.99, for pro plan is \$49.99, and for student plan is \$9.99.

We are interested in which factors may influence gym-goers' subscription plan type. 

## EDA

### Subcription Plan by Gender

```{r}
gender_plan_p =
  users_data |> 
  ggplot(aes(x = subscription_plan, fill = gender)) +
  geom_bar(position = "dodge", width = .6) +
  labs(
    title = "Subscription Plan by Gender", 
    x = "Subscription Plan", 
    y = "Count")

ggplotly(gender_plan_p)
```

The number of gym-goers choosing the pro subscription plan is similar to the number choosing the basic subscription plan, with slightly more men than women.

### Subscription Plan by Cities Where Users Live

```{r}
city_plan_p =
  users_data |> 
  ggplot(aes(x = user_location, fill = subscription_plan)) +
  geom_bar(position = "dodge", width = .7) +
  labs(
    title = "Subscription Plan by Cities Where Gym-goers Live", 
    x = "City", 
    y = "Count") +
  coord_flip()

ggplotly(city_plan_p)

city_plan =
  users_data |> 
  group_by(user_location, subscription_plan) |> 
  summarize(
    frequency = n()
  ) |> 
  pivot_wider(
    names_from = subscription_plan,
    names_prefix = "plan_",
    values_from = frequency
  )

city_plan |> 
  kable(digits = 3, caption = "Subscription Plan by Cities Where Gym-goers Live")
```

Atlanta has the most gym-goers choosing pro subscription plan and student plan, and Las Vegas has the most gym-goers choosing basic subscription plan.

### Distribution of Age across Subscription Plans

```{r}
age_plan_p =
  users_data |> 
  ggplot(aes(x = age, fill = subscription_plan)) +
  geom_density(linewidth = .2) + 
  facet_grid(subscription_plan ~ .) +
  labs(
    title = "Distribution of Age across Subscription Plans",
    x = "Age",
    y = "Density"
  )

ggplotly(age_plan_p)

age_plan = 
  users_data |> 
  group_by(subscription_plan) |> 
  summarize(
    age.min = min(age),
    age.max = max(age),
    age.mean = round(mean(age), 0),
  )
```

The age distribution of gym-goers is relatively even across different subscription plans. The youngest age is `r age_plan |> pull(age.min) |> min()` years old, the oldest is `r age_plan |> pull(age.max) |> max()` years old, and the mean age is `r age_plan |> pull(age.mean) |> mean()` years old.

## Statistical analysis

In this data analysis, we employ a logistic regression model to assess the relationship between the gym-goers' subscription plans and the average duration of their workouts, as well as average calories burned and cities where they live.

### Subscription Plan and Average duration

```{r}
factors_plan =
  final |> 
  filter(subscription_plan != "Student") |> 
  mutate(
    plan.pro = ifelse(subscription_plan == "Pro", 1, 0)
  ) |> 
  group_by(user_id, user_location, plan.pro) |> 
  summarize(
    calories.mean = mean(calories_burned),
    duration.mean = mean(workout_duration)
  )

fit_calories = glm(plan.pro ~ calories.mean, 
                   data = factors_plan, family = binomial())

fit_calories |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |> 
  kable(digits = 3)
```

### Subscription Plan and calories burned

```{r}
fit_duration = glm(plan.pro ~ duration.mean, 
                  data = factors_plan, family = binomial())

fit_duration |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |> 
  kable(digits = 3)
```

### Subscription Plan and cities

```{r}
fit_cities = glm(plan.pro ~ user_location, 
                    data = factors_plan, family = binomial())

fit_cities |> 
  broom::tidy() |> 
  mutate(
    term = sub("user_location", "", term)) |>
  select(user_location = term, estimate, p.value) |> 
  kable(digits = 3)
```

Overall, Las Vegas is the only city that shows a statistically significant negative impact on subscription plan (p < .05), while the effects of other cities are not significant.

Similarly, we are also interested in the relationship between subscription plan and gym types.

### Subscription Plan and Gym Types

```{r}
frequency_plan = 
  final |> 
  group_by(user_id, gym_type, subscription_plan) |> 
  summarize(
    frequency = n()
  )

frequency_plan |> 
  ggplot(aes(x = gym_type, y = frequency, fill = subscription_plan)) +
  geom_boxplot() + 
  labs(
    title = "Frequency of visits by Gym Types",
    x = "Age",
    y = "Density"
  )

frequency_fit = 
  frequency_plan |>
  lm(frequency ~ gym_type, data = _) |> 
  broom::tidy() |> 
  mutate(
    term = sub("gym_type", "", term)) |>
  select(gym_type = term, estimate, p.value) 

frequency_fit |> 
  kable(digits = 3)
```

The results of the linear regression model suggest that gym type significantly influences the frequency of gym visits (p < .05). 

On average, gym-goers visited the budget gyms `r frequency_fit$estimate[1] |> round(0)` times, premium gyms `r frequency_fit$estimate[1] |> round(0) + frequency_fit$estimate[2] |> round(0)` times, and standard gyms `r frequency_fit$estimate[1] |> round(0) + frequency_fit$estimate[3] |> round(0)` times in 2023.

## Conclusion

- We do not observe any difference in users' subscription plans by their gender, age, average duration of workouts and average calories burned. 

- Las Vegas has a statistically significant impact on subscription plan.

- Gym type plays a crucial role in determining how often users visit the gym.